---
title: "Math Scores and Science Scores by Gender"
output:
  html_document:
    fig_height: 3
    fig_width: 5
editor_options: 
  chunk_output_type: inline
---

Introduction
===========
Is there a relationship between a student’s math score and their science score? 
Is this association the same for female score as it is for male scores?
The goal of this analysis was to determine if there was an association between
a student's math score and their science score.  Further, this was used
to determine if this association was consistent across both genders. The 
standardized test scores from 200 students were used as data for the analysis.


```{r, include=FALSE}
library(tidyverse)

hsb2 <- read_csv("hsb2.csv")

```

Analysis
========

Math and Science Association
---------------------------
The correlation and 95% confidence interval of the math - science associations 
is included as descriptive statistics below. The scatterplot and regression line 
are included as a visual representation of the association between math and 
science scores of all students in the data set.  
```{r, echo=FALSE}

hsb2$female <- factor(hsb2$female, levels = c(0:1), 
                      labels = c("Male", "Female"))


male <- subset(hsb2, hsb2$female == "Male")
female <- subset(hsb2, hsb2$female == "Female")

ggplot(data = hsb2, mapping = aes(x = math, y = science)) +
  geom_point(male, mapping = aes(x = math, y = science), color = "Blue") +
  geom_point(female, mapping = aes(x = math, y = science), color = "Pink") +
  geom_text(aes(70, 30, label = "cor = 0.63")) +
  stat_smooth(method = "lm", fill = "grey") +
  theme_classic() +
  labs( x = "Math Score",
        y = "Science Score",
        title = "Math Scores Relation to Science Scores")


```



Math and Science by Gender
--------------------------

This data was then broken down by gender.  The same statistical analysis
was used to determine if a student's gender had an impact on the relationship 
between their math and science grades. 

The correlation and 95% confidence interval of the math - science associations 
for males and females separately is included as descriptive statistics below. 
A visual representation of the correlation for each gender are included below.   
```{r, echo=FALSE}


ggplot(data = male, mapping = aes(x = math, y = science)) +
  geom_point(color = "Blue") +
  geom_text(aes(70, 30, label = "cor = 0.61")) +
  stat_smooth(method = "lm", fill = "grey") +
  theme_classic() +
  labs( x = "Math Score",
        y = "Science Score",
        title = "Math Score Relation to Science Score in Males")


ggplot(data = female, mapping = aes(x = math, y = science)) +
  geom_point(color = "Pink") +
  geom_text(aes(70, 30, label = "cor = 0.66")) +
  stat_smooth(method = "lm", fill = "grey") +
  theme_classic() +
  labs( x = "Math Score",
        y = "Science Score",
        title = "Math Score Relation to Science Score in Females")



```


Conclusion
==========

The correlation coefficient for all students was 0.63. There is a moderately 
positive correlation between each student's math and science scores.  This means 
that students who earn high scores in math are likely to earn high scores in 
science.  When the data was broken down by gender, the correlation coefficient 
for females was 0.66.  This was higher than the correlation coefficient
for male students, which was 0.61. The analyses provided the 95% confidence
intervals for all students in the data set, and for the male and female students
separetely. 
 0.54 < all students < 0.71  &  0.46 < male < 0.73  &  0.53 < female < 0.75.  
The association between math scores and science scores is stronger for female 
students than for all students in the data and for just the male students.

